---
title: "Case Study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(broom)
library(stringr)
library(knitr)
library(nnet)
library(ggplot2)
```

## Introduction

### The Data

```{r}
data <- read.csv("/cloud/project/data/data-train.csv")
```


### Goals

Prediction: For a new parameter setting of (Re, F r, St), predict its particle cluster volume distribution in terms of its four raw moments.

Inference: Investigate and interpret how each parameter affects the probability distribution for particle cluster volumes

## Exploratory Data Analysis

First, we will explore the data to ensure it is fit for modelling and determine inital transformations needed of the data, and which model we see would best fit the data.

```{r}
names(data)
summary(data)

```

### Histograms

```{r}
hist(data$R_moment_1)
hist(data$R_moment_2)
hist(data$R_moment_3)
hist(data$R_moment_4)
hist(data$Fr, breaks=20)
hist(data$St, breaks=20)
hist(data$Re, breaks=20)


```

With these histograms its clear to see that each R_moment is heavily right skewed, since there are many rows of 0 in the data. In R_moment_3 and R_moment_4, the maximum values are extremely high whereas the medians are much smaller in comparison, which poses a problem to the analysis. We believe it is best then to apply a transformation to these variables in order to obtain more accurate analysis.


```{r}
hist(log(data$R_moment_1), breaks=20)
hist(log(data$R_moment_2), breaks=20)
hist(log(data$R_moment_3), breaks=20)
hist(log(data$R_moment_4), breaks=20)

```

Performing a log transformation on these variables created more normally distributed variables. While not perfectly normal, this is a big improvement to the non-transformed variables. From here on out, the log version of variables will be used and will be reflected as such in our interpretations and analysis. 

```{r}
pairs(data)
```
It appears that each R_moment variable has somewhat of a linear relationship with St. 

## Initial Modelling

We will fit a basic linear model onto each response variable.

```{r}
model1 <- lm(log(R_moment_1) ~ St + factor(Re) + factor(Fr), data=data)

summary(model1)

model2 <- lm(log(R_moment_2) ~ St + factor(Re) + factor(Fr), data=data)

summary(model2)

model3 <- lm(log(R_moment_3) ~ St + factor(Re) + factor(Fr), data=data)

summary(model3)

model4 <- lm(log(R_moment_4) ~ St + factor(Re) + factor(Fr), data=data)

summary(model4)
```


